{\TheTitle}

### 2017-07-11 greedy_hamiltonian_second_revision/chapters/4.sympmor.tex
second revision added

The greedy approach presented in Section \ref{Chap:Symo.PrSy:3} is an iterative method for construction of a symplectic basis. It avoids the evaluation of the full snapshot matrix, hence substantially reduces the computational cost in the offline stage of the symplectic model reduction. Also, by construction, it yields a orthosymplectic basis and therefore a well-conditioned basis.

@@

\begin{algorithm} 
\caption{The greedy algorithm for generation of a symplectic basis} \label{alg:SyMo:3}
{\bf Input:} Tolerated loss in the Hamiltonian $\delta$, parameter set $\Gamma_N = \{\omega_1,\dots,\omega_N\}$, initial condition $\mathbf z_0(\omega)$
\begin{enumerate}
\item $\omega^* \leftarrow \omega_1$
\item $e_1 \leftarrow \mathbf z_0(\omega^*)$
\item $A \leftarrow [e_1,-\mathbb J^T_{2n}e_1]$
\item $k \leftarrow 1$
\item \textbf{while} $\Delta H(\omega) > \delta$ for all $\omega \in \Omega_N$
\item \hspace{0.5cm} $w^* \leftarrow$ $\underset{\omega\in \Omega_N}{\text{argmax }}\Delta H(\omega)$
\item \hspace{0.5cm} Compute trajectory snapshots $S=\{ \mathbf z(t_i,\omega^*) | i = 1,\dots,M \}$
\item \hspace{0.5cm} $\mathbf z^* \leftarrow$ $\underset{s\in S}{\text{argmax }} \| s - AA^+s \|$
\item \hspace{0.5cm} Apply symplectic Gram-Schmidt on $\mathbf z^*$
\item \hspace{0.5cm} $e_{k+1} \leftarrow \mathbf z^*/ \| \mathbf  z^*\|$
\item \hspace{0.5cm} $A \leftarrow [e_1,\dots ,e_{k+1} , \mathbb J^T_{2n}e_1,\dots,\mathbb J^T_{2n}e_{k+1}]$
\item \hspace{0.5cm} $k \leftarrow k+1$
\item \textbf{end while}
\end{enumerate}
\vspace{0.5cm}
{\bf Output:} Symplectic basis $A$.
\end{algorithm}

@@

Suppose that we are given a compact subset $S$ of $\mathbb R^{2n}$. Our intention is to find a set of vectors $A=\{e_1,\dots,e_k,f_1,\dots,f_k\}$ such that $A$ forms {\edit an orthosymplectic} basis and any $s\in S$ is well approximated by elements of the subspace span$(A)$. The modified greedy method for generating basis vectors $e_i$ and $f_i$ is as follows. In the initial step we pick $e_1$ such that $\edit \|e_1\|_2 = \max_{s\in S} \|s\|_2$. Then define $f_1 = \mathbb{J}_{2n}^T e_1$. It is easy to check that the span of $A_2 = \{e_1,f_1\}$ is {\edit orthosymplectic}, so $A_2$ is the first subspace that approximates elements of $S$. In the $k$-th step of the greedy method, suppose we have a basis $A_{2k} = \{ e_1,\dots, e_k , f_1,\dots ,f_k \}$. We define $P_{2k}$ to be a symplectic projection operator that projects elements of $S$ onto span$(A_{2k})$ and define
\begin{equation} \label{eq:new1}
	\sigma_{2k}(s) := \|s-P_{2k}(s)\|_2,
\end{equation}
as the projection error. Moreover we denote by $\sigma_{2k}$ the maximum approximation error of $S$ using elements in span$(A_{2k})$ as
\begin{equation} \label{eq:new2}
	\sigma_{2k} := \max_{s\in S} \sigma_{2k}(s).
\end{equation}
The next set of basis vectors in the greedy selection are
\begin{equation} \label{eq:new3}
	e_{k+1} := \underset{s\in S}{\text{argmax }}\sigma_{2k}(s), \quad f_{k+1} := \mathbb{J}^T e_{k+1}.
\end{equation}
We emphasisze that the sequence of basis vectors generated by the greedy is generally not unique.

@@

We start by symplectifying the vectors provided by the greedy algorithm as
\begin{equation} \label{eq:new6}
\begin{aligned}
	& \xi_1 = e_i, & \bar{\xi}_1 = \mathbb{J}_{2n}^T \xi_1, &\\
	& \xi_i = e_i - P_{2(i-1)} (e_i), & \bar{\xi}_i = \mathbb{J}_{2n}^T, \xi_i &\quad i = 2,3,\dots
\end{aligned}
\end{equation}
The projection of a vector $s\in S$ onto span$(A_{2k})$ can be written using the symplectic basis as
\begin{equation} \label{eq:new7}
	P_{2k}(s) = \sum_{i=1}^k \left( \alpha_i(s) \xi_i + \bar{\alpha}_i(s) \bar{\xi}_i \right),
\end{equation}
where $\alpha_i(s)$ and $\bar{\alpha}_i(s)$ for $i=1,\dots,k$ are the expansion coefficients
\begin{equation} \label{eq:new8}
	\alpha_i(s) = - \frac{\Omega(\bar{\xi}_i,s)}{\Omega(\xi_i,\bar{\xi}_i)}, \quad \bar{\alpha}_i(s) = \frac{\Omega(\xi_i,s)}{\Omega(\xi_i,\bar{\xi}_i)},
\end{equation}
for any $s\in S$. Since $\bar{\xi}_i$ is symplectic to span$(A_{2(k-1)})$ we have
\begin{equation} \label{eq:new9}
\begin{aligned}
	|\alpha_i(s)| = \frac{|\Omega(\bar{\xi}_i,s)|}{|\Omega(\xi_i,\bar{\xi}_i)|} = \frac{|\Omega( \bar{\xi}_i, s - P_{2(k-1)}(s))|}{|\Omega(\xi_i,\bar{\xi}_i)|}  &\leq \frac{\|\bar{\xi}_i\|_2 \| s - P_{2(k-1)}(s) \|_2}{ \|\xi_i\|_2 \|\bar{\xi}_i\|_2 } \\
	&= \frac{\| s - P_{2(k-1)}(s) \|_2}{\| e_i - P_{2(k-1)}(e_i) \|_2} \leq 1.
\end{aligned}
\end{equation}
Here, we use the fact that $\edit |\Omega(\xi_i,\bar{\xi}_i)| = \| \xi_i \|^2_2 = \|\bar{\xi}_i\|^2_2$ with the last inequality following from the greedy algorithm which maximizes $e_i$. Similarly we deduce that $|\bar{\alpha}_i(s)|\leq 1$.

@@

Define $c = \kappa_j^{-1} \sum_{i=1,i\neq j}^{k+1} \kappa_i \xi_i + \kappa_j^{-1}\sum_{i=1}^{k+1} \kappa_{i+k+1} \bar{\xi}_i$ and note that $\mathbb{J}_{2n}^T c$ is symplectic to $\xi_j$. We recover

@@

\subsection{Symplectic Discrete Empirical Interpolation Method (SDEIM)} Consider the Hamiltonian system (\ref{eq:Hasy:7}) and its reduced system (\ref{eq:SyMo:4}) equipped with a symplectic transformation $A$. One can split the Hamiltonian function $H = H_1 + H_2$ such that $\nabla H_1 = L\mathbf z$ and $\nabla H_2 = \mathbf g(\mathbf z)$, where $L$ is a constant matrix in $\mathbb R^{n\times n}$ and $\mathbf g$ is a nonlinear function. The reduced system takes the form
\begin{equation} \label{eq:new24}
	\frac{d}{dt} \mathbf y = \underbrace{A^+ \mathbb J_{2n} L A}_{\tilde L} \mathbf y + A^+ \mathbb J_{2n} \mathbf g(A\mathbf y)
\end{equation}
As discussed in Section \ref{chap:MoOr.DEIM:1}, the complexity of evaluating the nonlinear term still depends on $n$, the size of the original system. To overcome this computational bottleneck we use the DEIM approximation for evaluating the nonlinear function $\mathbf g$ as
\begin{equation} \label{eq:new25}
	\frac{d}{dt} \mathbf y = \tilde L \mathbf y + \underbrace{ A^+ \mathbb J_{2n} V (P^TV)^{-1} P^T \mathbf g(A\mathbf y) }_{\tilde N(\mathbf y)}
\end{equation}
For a general choice of $V$ the system (\ref{eq:new25}) is not guaranteed to be a Hamiltonian system, impacting long time accuracy and stability. However, we can guarantee that (\ref{eq:new25}) is a Hamiltonian system by choosing $V=(A^+)^T$. To {\edit see this}, we note that the system (\ref{eq:new25}) is a Hamiltonian system if and only if $\tilde N(\mathbf y) = \mathbb J_{2k} \nabla_{\mathbf y} \mathbf g(\mathbf y)$. Also we have 
\begin{equation} \label{eq:new26}
	\mathbf g(A\mathbf y) = \nabla_{\mathbf z} H_2(\mathbf z) = (A^+)^T \nabla_{\mathbf y} H_2(A \mathbf y),
\end{equation}
where the chain rule is used for the second equality. Substituting this into $\tilde N$ we obtain
\begin{equation} \label{eq:new27}
	\tilde N(\mathbf y)= A^+ \mathbb J_{2n} V (P^TV)^{-1} P^T  (A^+)^T \nabla_{\mathbf y} H_2(A \mathbf y).
\end{equation}
Taking $V = (A^+)^T$ yields
\begin{equation} \label{eq:new28}
	\tilde N(\mathbf y) = A^+ \mathbb J_{2n}(A^+)^T \nabla_{\mathbf y} H_2(A \mathbf y) = \mathbb J_{2k} \nabla_{\mathbf y} H_2(A \mathbf y),
\end{equation}
since $(A^+)^T$ is a symplectic matrix. Hence, $V = (A^+)^T$ is a sufficient condition for (\ref{eq:new25}) to {\edit be Hamiltonian}.

@@

When using an implicit time integration scheme we face inefficiencies when evaluating the Jacobian of nonlinear terms, as discussed in Section \ref{chap:MoOr.DEIM:1}. We recall that the key to fast approximation of the Jacobian is that the interpolating index matrix $P$, obtained in the DEIM approximation, commutes with the nonlinear function. Nonlinear terms in Hamiltonian systems often take the from
\begin{equation}
	\mathbf g (\mathbf z) = \mathbf g (\mathbf q,\mathbf p) = 
	\begin{pmatrix}
		g_1(q_1,p_1) \\
		g_2(q_2,p_2) \\
		\vdots \\
		g_{2n}(q_{n},p_{n})
	\end{pmatrix}.
\end{equation}
Thus, the interpolating index matrix, obtained by Algorithm \ref{alg:MoOr:1} does not necessarily commute with the function $\mathbf g$. To overcome this, when index $\mathfrak p_i$ with $\mathfrak p_i\leq n$ or $\mathfrak p_i>n$ is chosen in Algorithm \ref{alg:MoOr:1} we also include $\mathfrak p_i + n$ or $\mathfrak p_i-n$, respectively. {\edit Simple calculations verifies that $\mathbf g$ and $P$ commute.}

>>>>

The greedy approach presented in Section \ref{Chap:Symo.PrSy:3} is an iterative method for construction of a symplectic basis. It avoids the evaluation of the full snapshot matrix, hence substantially reduces the computational cost in the offline stage of the symplectic model reduction. Also, by construction, it yields an orthosymplectic basis and therefore a well-conditioned basis.

@@

\begin{algorithm} 
\caption{The greedy algorithm for generation of a symplectic basis} \label{alg:SyMo:3}
{\bf Input:} Tolerated loss in the Hamiltonian $\delta$, parameter set $\Gamma_N = \{\omega_1,\dots,\omega_N\}$, initial condition $\mathbf z_0(\omega)$
\begin{enumerate}
\item $\omega^* \leftarrow \omega_1$
\item $e_1 \leftarrow \mathbf z_0(\omega^*)$
\item $A \leftarrow [e_1,\mathbb J^T_{2n}e_1]$
\item $k \leftarrow 1$
\item \textbf{while} $\Delta H(\omega) > \delta$ for all $\omega \in \Gamma_N$
\item \hspace{0.5cm} $\omega^* \leftarrow$ $\underset{\omega\in \Gamma_N}{\text{argmax }}\Delta H(\omega)$
\item \hspace{0.5cm} Compute trajectory snapshots $S=\{ \mathbf z(t_i,\omega^*) | i = 1,\dots,M \}$
\item \hspace{0.5cm} $\mathbf z^* \leftarrow$ $\underset{s\in S}{\text{argmax }} \| s - AA^+s \|$
\item \hspace{0.5cm} Apply symplectic Gram-Schmidt on $\mathbf z^*$
\item \hspace{0.5cm} $e_{k+1} \leftarrow \mathbf z^*/ \| \mathbf  z^*\|$
\item \hspace{0.5cm} $A \leftarrow [e_1,\dots ,e_{k+1} , \mathbb J^T_{2n}e_1,\dots,\mathbb J^T_{2n}e_{k+1}]$
\item \hspace{0.5cm} $k \leftarrow k+1$
\item \textbf{end while}
\end{enumerate}
\vspace{0.5cm}
{\bf Output:} Symplectic basis $A$.
\end{algorithm}

@@

Suppose that we are given a compact subset $S$ of $\mathbb R^{2n}$. Our intention is to find a set of vectors $A=\{e_1,\dots,e_k,f_1,\dots,f_k\}$ such that $A$ forms {\edit an orthosymplectic} basis and any $s\in S$ is well approximated by elements of the subspace span$(A)$. The modified greedy method for generating basis vectors $e_i$ and $f_i$ is as follows. In the initial step we pick $e_1$ such that $\edit \|e_1\|_2 = \max_{s\in S} \|s\|_2$. Then define $f_1 = \mathbb{J}_{2n}^T e_1$. It is easy to check that the span of $A_2 = \{e_1,f_1\}$ is {\edit orthosymplectic}, so $A_2$ is the first subspace that approximates elements of $S$. In the $k$-th step of the greedy method, suppose we have a basis $A_{2k} = \{ e_1,\dots, e_k , f_1,\dots ,f_k \}$. We define $P_{2k}$ to be a symplectic projection operator that projects elements of $S$ onto span$(A_{2k})$ and define
\begin{equation} \label{eq:new1}
	\sigma_{2k}(s) := \|s-P_{2k}(s)\|_2,
\end{equation}
as the projection error. Moreover we denote by $\sigma_{2k}$ the maximum approximation error of $S$ using elements in span$(A_{2k})$ as
\begin{equation} \label{eq:new2}
	\sigma_{2k} := \max_{s\in S} \sigma_{2k}(s).
\end{equation}
The next set of basis vectors in the greedy selection are
\begin{equation} \label{eq:new3}
	e_{k+1} := \underset{s\in S}{\text{argmax }}\sigma_{2k}(s), \quad f_{k+1} := \mathbb{J}_{2n}^T e_{k+1}.
\end{equation}
We emphasisze that the sequence of basis vectors generated by the greedy is generally not unique.

@@

We start by $\edit \mathbb J_{2n}$-orthogonalizing the vectors provided by the greedy algorithm as
\begin{equation} \label{eq:new6}
\begin{aligned}
	& \xi_1 = e_i, & \bar{\xi}_1 = \mathbb{J}_{2n}^T \xi_1, &\\
	& \xi_i = e_i - P_{2(i-1)} (e_i), & \bar{\xi}_i = \mathbb{J}_{2n}^T, \xi_i &\quad i = 2,3,\dots
\end{aligned}
\end{equation}
The projection of a vector $s\in S$ onto span$(A_{2k})$ can be written using the symplectic basis as
\begin{equation} \label{eq:new7}
	P_{2k}(s) = \sum_{i=1}^k \left( \alpha_i(s) \xi_i + \bar{\alpha}_i(s) \bar{\xi}_i \right),
\end{equation}
where $\alpha_i(s)$ and $\bar{\alpha}_i(s)$ for $i=1,\dots,k$ are the expansion coefficients
\begin{equation} \label{eq:new8}
	\alpha_i(s) = - \frac{\Omega(\bar{\xi}_i,s)}{\Omega(\xi_i,\bar{\xi}_i)}, \quad \bar{\alpha}_i(s) = \frac{\Omega(\xi_i,s)}{\Omega(\xi_i,\bar{\xi}_i)},
\end{equation}
for any $s\in S$. Since $\bar{\xi}_i$ is {\edit $\mathbb J_{2n}$-orthogonal} to the span$(A_{2(k-1)})$ we have
\begin{equation} \label{eq:new9}
\begin{aligned}
	|\alpha_i(s)| = \frac{|\Omega(\bar{\xi}_i,s)|}{|\Omega(\xi_i,\bar{\xi}_i)|} = \frac{|\Omega( \bar{\xi}_i, s - P_{2(k-1)}(s))|}{|\Omega(\xi_i,\bar{\xi}_i)|}  &\leq \frac{\|\bar{\xi}_i\|_2 \| s - P_{2(k-1)}(s) \|_2}{ \|\xi_i\|_2 \|\bar{\xi}_i\|_2 } \\
	&= \frac{\| s - P_{2(k-1)}(s) \|_2}{\| e_i - P_{2(k-1)}(e_i) \|_2} \leq 1.
\end{aligned}
\end{equation}
Here, we use the fact that $\edit |\Omega(\xi_i,\bar{\xi}_i)| = \| \xi_i \|^2_2 = \|\bar{\xi}_i\|^2_2$ with the last inequality following from the greedy algorithm which maximizes $e_i$. Similarly we deduce that $|\bar{\alpha}_i(s)|\leq 1$.

@@

Define $c = \kappa_j^{-1} \sum_{i=1,i\neq j}^{k+1} \kappa_i \xi_i + \kappa_j^{-1}\sum_{i=1}^{k+1} \kappa_{i+k+1} \bar{\xi}_i$. Using that $\mathbb{J}_{2n}^T c$ is {\edit $\mathbb J_{2n}$-orthogonal} to $\xi_j$ we recover

@@

\subsection{Symplectic Discrete Empirical Interpolation Method (SDEIM)} Consider the Hamiltonian system (\ref{eq:Hasy:7}) and its reduced system (\ref{eq:SyMo:4}) equipped with a symplectic transformation $A$. One can split the Hamiltonian function $H = H_1 + H_2$ such that $\nabla H_1 = L\mathbf z$ and $\nabla H_2 = \mathbf g(\mathbf z)$, where $L$ is a constant matrix in $\edit \mathbb R^{2n\times 2n}$ and $\mathbf g$ is a nonlinear function. The reduced system takes the form
\begin{equation} \label{eq:new24}
	\frac{d}{dt} \mathbf y = \underbrace{A^+ \mathbb J_{2n} L A}_{\tilde L} \mathbf y + A^+ \mathbb J_{2n} \mathbf g(A\mathbf y)
\end{equation}
As discussed in Section \ref{chap:MoOr.DEIM:1}, the complexity of evaluating the nonlinear term still depends on $n$, the size of the original system. To overcome this computational bottleneck we use the DEIM approximation for evaluating the nonlinear function $\mathbf g$ as
\begin{equation} \label{eq:new25}
	\frac{d}{dt} \mathbf y = \tilde L \mathbf y + \underbrace{ A^+ \mathbb J_{2n} V (P^TV)^{-1} P^T \mathbf g(A\mathbf y) }_{\tilde N(\mathbf y)}
\end{equation}
For a general choice of $V$ the system (\ref{eq:new25}) is not guaranteed to be a Hamiltonian system, impacting long time accuracy and stability. However, we can guarantee that (\ref{eq:new25}) is a Hamiltonian system by choosing $V=(A^+)^T$. To {\edit see this}, we note that the system (\ref{eq:new25}) is a Hamiltonian system if and only if $\tilde N(\mathbf y) = \mathbb J_{2k} \nabla_{\mathbf y} \mathbf g(\mathbf y)$. Also we have 
\begin{equation} \label{eq:new26}
	\mathbf g(A\mathbf y) = \nabla_{\mathbf z} H_2(\mathbf z) = (A^+)^T \nabla_{\mathbf y} H_2(A \mathbf y),
\end{equation}
where the chain rule is used for the second equality. Substituting this into $\tilde N$ we obtain
\begin{equation} \label{eq:new27}
	\tilde N(\mathbf y)= A^+ \mathbb J_{2n} V (P^TV)^{-1} P^T  (A^+)^T \nabla_{\mathbf y} H_2(A \mathbf y).
\end{equation}
Taking $V = (A^+)^T$ yields
\begin{equation} \label{eq:new28}
	\tilde N(\mathbf y) = A^+ \mathbb J_{2n}(A^+)^T \nabla_{\mathbf y} H_2(A \mathbf y) = \mathbb J_{2k} \nabla_{\mathbf y} H_2(A \mathbf y),
\end{equation}
since $(A^+)^T$ is a symplectic matrix. Hence, $V = (A^+)^T$ is a sufficient condition for (\ref{eq:new25}) to {\edit be Hamiltonian}.

@@

When using an implicit time integration scheme we face inefficiencies when evaluating the Jacobian of nonlinear terms, as discussed in Section \ref{chap:MoOr.DEIM:1}. We recall that the key to fast approximation of the Jacobian is that the interpolating index matrix $P$, obtained in the DEIM approximation, commutes with the nonlinear function. Nonlinear terms in Hamiltonian systems often take the from
\begin{equation}
	\mathbf g (\mathbf z) = \mathbf g (\mathbf q,\mathbf p) = 
	\begin{pmatrix}
		g_1(q_1,p_1) \\
		g_2(q_2,p_2) \\
		\vdots \\
		g_{2n}(q_{n},p_{n})
	\end{pmatrix}.
\end{equation}
Thus, the interpolating index matrix, obtained by Algorithm \ref{alg:MoOr:1} does not necessarily commute with the function $\mathbf g$. To overcome this, when index $\mathfrak p_i$ with $\mathfrak p_i\leq n$ or $\mathfrak p_i>n$ is chosen in Algorithm \ref{alg:MoOr:1} we also include $\mathfrak p_i + n$ or $\mathfrak p_i-n$, respectively. {\edit Simple calculations verifies that $\mathbf g$ and $P$ then commute.}

### 2017-07-11 greedy_hamiltonian_second_revision/chapters/5.numres.tex
second revision added

\section{Numerical Results} \label{chap:NuRe:1}
In this section, we illustrate the performance of the greedy generation of a symplectic basis. The parametric linear wave equation is considered to compare SVD based methods with the greedy method. The nonlinear model order reduction using the combination of DIEM and the symplectic basis is then illustrated by considering the parametric nonlinear Schr\"odinger equation. {\edit Finally we discuss the numerical convergence of the greedy method introduced in Algorithm \ref{alg:SyMo:3}.}

>>>>

\section{Numerical Results} \label{chap:NuRe:1}
In this section, we illustrate the performance of the greedy generation of a symplectic basis. The parametric linear wave equation is considered to compare SVD based methods with the greedy method. {\blue The symplectic model reduction of nonlinear systems} is then illustrated by considering the parametric nonlinear Schr\"odinger equation. {\edit Finally we discuss the numerical convergence of the greedy method introduced in Algorithm \ref{alg:SyMo:3}.}

### 2017-07-11 greedy_hamiltonian_second_revision/chapters/6.conclusion.tex
second revision added

In this paper, we present a greedy approach for the construction of a reduced system that preserves the geometric structure of Hamiltonian systems. An iteration of the greedy method comprises searching the parameter space using the error in the Hamiltonian, to find the best basis vectors that increase the overall accuracy of the reduced basis. We argued that for a compact subset with exponentially small Kolmogorov $n$-width we recover exponentially fast convergence of the greedy algorithm. For fast approximation of nonlinear terms, the basis obtained by the greedy was combined with the symplectic DEIM to construct a Hamiltonian reduced system.

@@

The numerical results demonstrated that the greedy method can save substantial computational cost in the offline stage as compared to alternative SVD-based techniques. Also since the reduced system obtained by the greedy method is Hamiltonian, the greedy method yields a stable reduced system. Symplectic DEIM also effectively reduced computational cost of approximating nonlinear terms while preserving stability and symplectic structure. Hence, the greedy method is an efficient model reduction technique that provides an accurate and stable reduced system for large-scale parametric Hamiltonian systems.

>>>>

In this paper, we present a greedy approach for the construction of a reduced system that preserves the geometric structure of Hamiltonian systems. An iteration of the greedy method comprises searching the parameter space using the error in the Hamiltonian, to find the best basis vectors that increase the overall accuracy of the reduced basis. We argue that for a compact subset with exponentially small Kolmogorov $n$-width we recover exponentially fast convergence of the greedy algorithm. For fast approximation of nonlinear terms, the basis obtained by the greedy was combined with a symplectic DEIM to construct {\edit a reduced system with a Hamiltonian that is arbitrary close to the Hamiltonian of the original system.}

@@

The numerical results demonstrate that the greedy method can save substantial computational cost in the offline stage as compared to alternative SVD-based techniques. Also since the reduced system obtained by the greedy method is Hamiltonian, the greedy method yields a stable reduced system. Symplectic DEIM effectively reduces computational cost of approximating nonlinear terms while preserving stability and symplectic structure. Hence, the greedy method is an efficient model reduction technique that provides an accurate and stable reduced system for large-scale parametric Hamiltonian systems.

### 2017-07-11 greedy_hamiltonian_second_revision/paper.tex
second revision added

\newcommand{\edit}{\color{black}}

@@

% REQUIRED
\begin{abstract}
While reduced-order models (ROMs) {\edit have been} popular for {\edit efficiently} solving large systems of differential equations, the stability of reduced models over long-time integration {\edit is of present challenges}. We present a greedy approach for ROM generation of parametric Hamiltonian systems {\edit that} captures the symplectic structure of Hamiltonian systems to ensure stability of the reduced model. Through the greedy selection of basis vectors, {\edit two new vectors are added at each iteration to the linear vector space} to increase the accuracy of the {\edit reduced} basis. We {\edit use} the error in the Hamiltonian due to model reduction as an error indicator to search the parameter space and identify the next best basis vectors. Under natural assumptions on the set of all solutions of the Hamiltonian system under variation of the parameters, we show that the greedy algorithm converges with exponential rate. Moreover, we demonstrate that combining the greedy basis with the discrete empirical interpolation method also preserves the symplectic structure. This enables the reduction of the computational cost for nonlinear Hamiltonian systems. The efficiency, accuracy, and stability of this model reduction technique is illustrated through simulations of the parametric wave equation and the parametric Schr\"odinger equation.
\end{abstract}

>>>>

\newcommand{\edit}{\color{black}}
\newcommand{\blue}{\color{blue}}

@@

% REQUIRED
\begin{abstract}
While reduced-order models (ROMs) {\edit have been} popular for {\edit efficiently} solving large systems of differential equations, the stability of reduced models over long-time integration {\edit is of present challenges}. We present a greedy approach for {\blue a} ROM generation of parametric Hamiltonian systems {\edit that} captures the symplectic structure of Hamiltonian systems to ensure stability of the reduced model. Through the greedy selection of basis vectors, {\edit two new vectors are added at each iteration to the linear vector space} to increase the accuracy of the {\edit reduced} basis. We {\edit use} the error in the Hamiltonian due to model reduction as an error indicator to search the parameter space and identify the next best basis vectors. Under natural assumptions on the set of all solutions of the Hamiltonian system under variation of the parameters, we show that the greedy algorithm converges with exponential rate. Moreover, we demonstrate that combining the greedy basis with the discrete empirical interpolation method also preserves the symplectic structure. This enables the reduction of the computational cost for nonlinear Hamiltonian systems. The efficiency, accuracy, and stability of this model reduction technique is illustrated through simulations of the parametric wave equation and the parametric Schr\"odinger equation.
\end{abstract}
